HTLSDstack

### 2025-05-25 Code/R/Research/HTLSDstack.Rmd
Add real parameter estimation case for htlsd algs

```{r setup, include=FALSE}
library(Rssa)
library(rTensor)
library(ggplot2)
library(dplyr)
library(tidyr)
library(future.apply)
library(progressr)

@@

```{r}
rrmse <- function(true, pred, k) {
  sqrt(sum((pred - true[k])^2) / length(pred)) * 100 / (0.5 * abs(true[1] - true[2]))
}

@@

  est <- list(
    esprit = parestimate(ssa(x, L = L_esprit), groups = groups_esprit, drop = FALSE),
    htlsd = HTLSDstack(
      x,
      D = D,
      L = L_htlsd,
      groups = groups_htlsd,
      method = "base"
    ),
    tens = HTLSDstack(
      x,
      D = D,
      L = L_tens,
      groups = groups_tens,
      method = "hooi",
      r3 = r3,
      est_dim = est_dim,
      status = FALSE
    )
  )

@@

# Example from [paper](https://doi.org/10.1002/cem.1212)
```{r}

# Example 1
A <- c(1, 1)
phi <- c(0, 0)
rate <- c(0, 0)
freq <- c(0.2, 0.205)
L_esprit <- 500
L_htlsd <- 50
L_tens <- 50
D <- 10
groups_esprit <- list(1:2)
groups_htlsd <- list(1:2)
groups_tens <- list(1:2)
r3 <- 2
est_dim <- 1
dt <- 0.1
N <- 1000
true_freq <- freq * dt

s <- colSums(A * exp(1i * phi) * exp((-rate + 1i * 2 * pi * freq) %o% seq(
  from = dt,
  length.out = N,
  by = dt
)))
```

```{r}
sd <- snr_to_sd(s, snr = 30)
R <- 100

estims <- replicate(R, compare_htls())
```

>>>>

```{r setup, include=FALSE}
library(svd)
library(gridExtra)
library(latex2exp)
library(Rssa)
library(rTensor)
library(ggplot2)
library(dplyr)
library(tidyr)
library(future.apply)
library(progressr)

@@

```{r}
rmse <- function(true, pred) {
  sqrt(mean((pred - true)^2))
}

rmse_mat <- function(true, pred) {
  apply(cbind(true, pred), 1, function(v) rmse(v[1], v[-1]))
}

rrmse_freq <- function(true, pred) {
  rmse_mat(true, pred) * 100 / (0.5 * abs(true[1] - true[2]))
}

@@

  est <- list(
    esprit = parestimate(ssa(x, L = L_esprit), groups = groups_esprit, drop = FALSE),
    htlsd = HTLSDstack(
      x,
      D = D,
      L = L_htlsd,
      groups = groups_htlsd,
      method = "base"
    ),
    tens1 = HTLSDstack(
      x,
      D = D,
      L = L_tens,
      groups = groups_tens,
      method = "hooi",
      r3 = r3_1,
      est_dim = est_dim,
      status = FALSE
    ),
    tens2 = HTLSDstack(
      x,
      D = D,
      L = L_tens,
      groups = groups_tens,
      method = "hooi",
      r3 = r3_2,
      est_dim = est_dim,
      status = FALSE
    )
  )

@@

# Examples from [paper](https://doi.org/10.1002/cem.1212)
```{r}

# Example 1
plot_name <- "complex_param_rmse"
A <- c(1, 1)
phi <- c(0, 0)
rate <- c(0, 0)
freq <- c(0.2, 0.205)
L_esprit <- 500
L_htlsd <- 50
L_tens <- 50
D <- 10
groups_esprit <- list(1:2)
groups_htlsd <- list(1:2)
groups_tens <- list(1:2)
r3_1 <- 2
r3_2 <- 1
est_dim <- 1
dt <- 0.1
N <- 999

s <- colSums(A * exp(1i * phi) * exp((-rate + 1i * 2 * pi * freq) %o% seq(
  from = 0,
  length.out = N,
  by = dt
)))
```

```{r warning=FALSE}
# sd <- snr_to_sd(s, snr = 35)
out_len <- 16
sds <- seq(0, 1.5, length.out = out_len)
rmse_df1 <- data.frame(sd = sds, esprit = numeric(out_len), HTLSDstack = numeric(out_len),
                      tens1 = numeric(out_len), tens2 = numeric(out_len))
rmse_df2 <- data.frame(sd = sds, esprit = numeric(out_len), HTLSDstack = numeric(out_len),
                      tens1 = numeric(out_len), tens2 = numeric(out_len))
R <- 100
for (i in seq(sds)) {
  print(i)
  sd <- sds[i]
  with(plan(multisession), {
    batch_estims <- function(R) {
      p <- progressor(R)
      future_replicate(R, {
        p()
        compare_htls()
      }, simplify = "array", future.seed = 5)
    }
    estims <- batch_estims(R)
  })

  freq_estims <- apply(estims[3:4, , ], 2:3, function(v)
    c(v[which.min(abs(v))], v[which.max(abs(v))]) / dt)
  freq_rmse <- apply(freq_estims, 2, rmse_mat, true = freq)
  rmse_df1[i,-1] <- freq_rmse[1,]
  rmse_df2[i,-1] <- freq_rmse[2,]
}

save(rmse_df1, rmse_df2, file = paste0("./Research/comp_results/htlsd_", plot_name, ".RData"))
```

```{r}
load(file = paste0("./Research/comp_results/htlsd_", plot_name, ".RData"))
sd_bound <- 0.75
p1 <- rmse_df1 |> filter(sd <= sd_bound) |> pivot_longer(cols = -sd,
                                                   names_to = "Method",
                                                   values_to = "RMSE") |>
  ggplot(aes(x = sd, y = RMSE)) +
  geom_line(aes(col = Method, lty = Method)) +
  scale_linetype_discrete(labels = c("ESPRIT", "HTLSDstack", "Tensor, r3 = 2", "Tesnor, r3 = 1")) +
  scale_color_discrete(labels = c("ESPRIT", "HTLSDstack", "Tensor, r3 = 2", "Tesnor, r3 = 1")) +
  ggtitle(TeX(sprintf("$\\nu = %5.3f$", freq[1])))
p2 <- rmse_df1 |> filter(sd >= sd_bound) |> pivot_longer(cols = -sd,
                                                    names_to = "Method",
                                                    values_to = "RMSE") |>
  ggplot(aes(x = sd, y = RMSE)) +
  geom_line(aes(col = Method, lty = Method)) +
  scale_linetype_discrete(labels = c("ESPRIT", "HTLSDstack", "Tensor, r3 = 2", "Tesnor, r3 = 1")) +
  scale_color_discrete(labels = c("ESPRIT", "HTLSDstack", "Tensor, r3 = 2", "Tesnor, r3 = 1")) +
  ggtitle(TeX(sprintf("$\\nu = %5.3f$", freq[1])))
pdf(paste0("./img/htlsd_", plot_name, "_1.pdf"), width = 9, height = 7)
grid.arrange(p1, p2)
dev.off()
grid.arrange(p1, p2)

p1 <- rmse_df2 |> filter(sd <= sd_bound) |> pivot_longer(cols = -sd, names_to = "Method", values_to = "RMSE") |>
  ggplot(aes(x = sd, y = RMSE)) + 
  geom_line(aes(col = Method, lty = Method)) +
  scale_linetype_discrete(labels = c("ESPRIT", "HTLSDstack", "Tensor, r3 = 2", "Tesnor, r3 = 1")) +
  scale_color_discrete(labels = c("ESPRIT", "HTLSDstack", "Tensor, r3 = 2", "Tesnor, r3 = 1")) +
  ggtitle(TeX(sprintf("$\\nu = %5.3f$", freq[2])))
p2 <- rmse_df2 |> filter(sd >= sd_bound) |> pivot_longer(cols = -sd, names_to = "Method", values_to = "RMSE") |>
  ggplot(aes(x = sd, y = RMSE)) + 
  geom_line(aes(col = Method, lty = Method)) +
  scale_linetype_discrete(labels = c("ESPRIT", "HTLSDstack", "Tensor, r3 = 2", "Tesnor, r3 = 1")) +
  scale_color_discrete(labels = c("ESPRIT", "HTLSDstack", "Tensor, r3 = 2", "Tesnor, r3 = 1")) +
  ggtitle(TeX(sprintf("$\\nu = %5.3f$", freq[2])))
pdf(paste0("./img/htlsd_", plot_name, "_2.pdf"), width = 9, height = 7)
grid.arrange(p1, p2)
dev.off()
grid.arrange(p1, p2)
```

# Real case with same parameters
```{r}
plot_name <- "real_param_rmse"
A <- c(1, 1)
phi <- c(0, 0)
rate <- c(0, 0)
freq <- c(0.2, 0.205)
L_esprit <- 500
L_htlsd <- 50
L_tens <- 50
D <- 10
groups_esprit <- list(1:4)
groups_htlsd <- list(1:4)
groups_tens <- list(1:4)
r3_1 <- 2
r3_2 <- 1
est_dim <- 1
dt <- 0.1
N <- 999

s <- colSums(A * exp(-rate) * cos(phi + 2 * pi * freq %o% seq(
  from = 0,
  length.out = N,
  by = dt
)))
```

```{r warning=FALSE}
# sd <- snr_to_sd(s, snr = 35)
out_len <- 16
sds <- seq(0, 1.5, length.out = out_len)
rmse_df11 <- data.frame(sd = sds, esprit = numeric(out_len), HTLSDstack = numeric(out_len),
                      tens1 = numeric(out_len), tens2 = numeric(out_len))
rmse_df12 <- data.frame(sd = sds, esprit = numeric(out_len), HTLSDstack = numeric(out_len),
                      tens1 = numeric(out_len), tens2 = numeric(out_len))
rmse_df21 <- data.frame(sd = sds, esprit = numeric(out_len), HTLSDstack = numeric(out_len),
                      tens1 = numeric(out_len), tens2 = numeric(out_len))
rmse_df22 <- data.frame(sd = sds, esprit = numeric(out_len), HTLSDstack = numeric(out_len),
                      tens1 = numeric(out_len), tens2 = numeric(out_len))
R <- 100
for (i in seq(sds)) {
  print(i)
  sd <- sds[i]
  with(plan(multisession), {
    batch_estims <- function(R) {
      p <- progressor(R)
      future_replicate(R, {
        p()
        compare_htls()
      }, simplify = "array", future.seed = 5)
    }
    estims <- batch_estims(R)
  })

  freq_estims <- apply(estims[5:8, , ], 2:3, function(v)
    v[order(v, abs(v))] / dt)
  freq_rmse <- apply(freq_estims, 2, rmse_mat, true = c(-rev(freq), freq))
  rmse_df11[i,-1] <- freq_rmse[1,]
  rmse_df12[i,-1] <- freq_rmse[2,]
  rmse_df21[i,-1] <- freq_rmse[3,]
  rmse_df22[i,-1] <- freq_rmse[4,]
}
save(rmse_df11, rmse_df12, rmse_df21, rmse_df22, file = paste0("./Research/comp_results/htlsd_", plot_name, ".RData"))
```

```{r}
load(file = paste0("./Research/comp_results/htlsd_", plot_name, ".RData"))
sd_bound <- 0.4
p1 <- rmse_df11 |> filter(sd <= sd_bound) |> pivot_longer(cols = -sd,
                                                   names_to = "Method",
                                                   values_to = "RMSE") |>
  ggplot(aes(x = sd, y = RMSE)) +
  geom_line(aes(col = Method, lty = Method)) +
  scale_linetype_discrete(labels = c("ESPRIT", "HTLSDstack", "Tensor, r3 = 2", "Tesnor, r3 = 1")) +
  scale_color_discrete(labels = c("ESPRIT", "HTLSDstack", "Tensor, r3 = 2", "Tesnor, r3 = 1")) +
  ggtitle(TeX(sprintf("$\\nu = %5.3f$", freq[1])))
p2 <- rmse_df11 |> filter(sd >= sd_bound) |> pivot_longer(cols = -sd,
                                                    names_to = "Method",
                                                    values_to = "RMSE") |>
  ggplot(aes(x = sd, y = RMSE)) +
  geom_line(aes(col = Method, lty = Method)) +
  scale_linetype_discrete(labels = c("ESPRIT", "HTLSDstack", "Tensor, r3 = 2", "Tesnor, r3 = 1")) +
  scale_color_discrete(labels = c("ESPRIT", "HTLSDstack", "Tensor, r3 = 2", "Tesnor, r3 = 1")) +
  ggtitle(TeX(sprintf("$\\nu = %5.3f$", freq[1])))
pdf(paste0("./img/htlsd_", plot_name, "_11.pdf"), width = 9, height = 7)
grid.arrange(p1, p2)
dev.off()
grid.arrange(p1, p2)

p1 <- rmse_df12 |> filter(sd <= sd_bound) |> pivot_longer(cols = -sd,
                                                   names_to = "Method",
                                                   values_to = "RMSE") |>
  ggplot(aes(x = sd, y = RMSE)) +
  geom_line(aes(col = Method, lty = Method)) +
  scale_linetype_discrete(labels = c("ESPRIT", "HTLSDstack", "Tensor, r3 = 2", "Tesnor, r3 = 1")) +
  scale_color_discrete(labels = c("ESPRIT", "HTLSDstack", "Tensor, r3 = 2", "Tesnor, r3 = 1")) +
  ggtitle(TeX(sprintf("$\\nu = %5.3f$", freq[1])))
p2 <- rmse_df12 |> filter(sd >= sd_bound) |> pivot_longer(cols = -sd,
                                                    names_to = "Method",
                                                    values_to = "RMSE") |>
  ggplot(aes(x = sd, y = RMSE)) +
  geom_line(aes(col = Method, lty = Method)) +
  scale_linetype_discrete(labels = c("ESPRIT", "HTLSDstack", "Tensor, r3 = 2", "Tesnor, r3 = 1")) +
  scale_color_discrete(labels = c("ESPRIT", "HTLSDstack", "Tensor, r3 = 2", "Tesnor, r3 = 1")) +
  ggtitle(TeX(sprintf("$\\nu = %5.3f$", freq[1])))
pdf(paste0("./img/htlsd_", plot_name, "_12.pdf"), width = 9, height = 7)
grid.arrange(p1, p2)
dev.off()
grid.arrange(p1, p2)

p1 <- rmse_df21 |> filter(sd <= sd_bound) |> pivot_longer(cols = -sd, names_to = "Method", values_to = "RMSE") |>
  ggplot(aes(x = sd, y = RMSE)) + 
  geom_line(aes(col = Method, lty = Method)) +
  scale_linetype_discrete(labels = c("ESPRIT", "HTLSDstack", "Tensor, r3 = 2", "Tesnor, r3 = 1")) +
  scale_color_discrete(labels = c("ESPRIT", "HTLSDstack", "Tensor, r3 = 2", "Tesnor, r3 = 1")) +
  ggtitle(TeX(sprintf("$\\nu = %5.3f$", freq[2])))
p2 <- rmse_df21 |> filter(sd >= sd_bound) |> pivot_longer(cols = -sd, names_to = "Method", values_to = "RMSE") |>
  ggplot(aes(x = sd, y = RMSE)) + 
  geom_line(aes(col = Method, lty = Method)) +
  scale_linetype_discrete(labels = c("ESPRIT", "HTLSDstack", "Tensor, r3 = 2", "Tesnor, r3 = 1")) +
  scale_color_discrete(labels = c("ESPRIT", "HTLSDstack", "Tensor, r3 = 2", "Tesnor, r3 = 1")) +
  ggtitle(TeX(sprintf("$\\nu = %5.3f$", freq[2])))
pdf(paste0("./img/htlsd_", plot_name, "_21.pdf"), width = 9, height = 7)
grid.arrange(p1, p2)
dev.off()
grid.arrange(p1, p2)

p1 <- rmse_df22 |> filter(sd <= sd_bound) |> pivot_longer(cols = -sd, names_to = "Method", values_to = "RMSE") |>
  ggplot(aes(x = sd, y = RMSE)) + 
  geom_line(aes(col = Method, lty = Method)) +
  scale_linetype_discrete(labels = c("ESPRIT", "HTLSDstack", "Tensor, r3 = 2", "Tesnor, r3 = 1")) +
  scale_color_discrete(labels = c("ESPRIT", "HTLSDstack", "Tensor, r3 = 2", "Tesnor, r3 = 1")) +
  ggtitle(TeX(sprintf("$\\nu = %5.3f$", freq[2])))
p2 <- rmse_df22 |> filter(sd >= sd_bound) |> pivot_longer(cols = -sd, names_to = "Method", values_to = "RMSE") |>
  ggplot(aes(x = sd, y = RMSE)) + 
  geom_line(aes(col = Method, lty = Method)) +
  scale_linetype_discrete(labels = c("ESPRIT", "HTLSDstack", "Tensor, r3 = 2", "Tesnor, r3 = 1")) +
  scale_color_discrete(labels = c("ESPRIT", "HTLSDstack", "Tensor, r3 = 2", "Tesnor, r3 = 1")) +
  ggtitle(TeX(sprintf("$\\nu = %5.3f$", freq[2])))
pdf(paste0("./img/htlsd_", plot_name, "_22.pdf"), width = 9, height = 7)
grid.arrange(p1, p2)
dev.off()
grid.arrange(p1, p2)
```
